---
title: "PA1_template.Rmd"
author: "wog48"
date: "Sunday, May 17, 2015"
output: html_document
---

## Loading and preprocessing the data
The raw data can be found in file activity.csv
```{r}
steps <- read.csv("activity.csv")
Sys.setlocale("LC_TIME", "English")
```
## What is mean total number of steps taken per day?
The following diagram shall show the distribution of the total number of steps done each day.
```{r}
stepsPerDay <- steps[!is.na(steps$steps),]
stepsPerDay <- aggregate(stepsPerDay$steps, by = list(stepsPerDay$date), FUN = sum)
names(stepsPerDay) <- c("Date","Steps")
hist(stepsPerDay$Steps, main = "Distribution of steps taken per day", xlab = "Steps per Day", breaks = 20)
```

## What is the average daily activity pattern?
The next diagram shall show how in average the activity is distribuited of a day.
```{r}
avgSteps <- steps[!is.na(steps$steps),]
avgSteps <- aggregate(avgSteps$steps, by = list(avgSteps$interval), FUN = mean)
convertTime <- function(timestamp){
                     if(nchar(timestamp) == 1) timestamp <- paste('000', timestamp, sep = "" )
                else if(nchar(timestamp) == 2) timestamp <- paste('00', timestamp, sep = "" )
                else if(nchar(timestamp) == 3) timestamp <- paste('0', timestamp, sep = "" )
                ##Need to use dot as separator instead of colon, as plot would not interpret e.g. 13:15 as n number.
                timestamp <- paste(substr(timestamp,start=1,stop=2),".",substr(timestamp,start = 3, stop = 4),sep="")
}
for(i in 1:length(avgSteps[,1])){
                avgSteps[i,1] <- convertTime(avgSteps[i,1])
}
plot(avgSteps$Group.1, avgSteps$x, xlab = "Five Minutes Intervalls [h]", ylab = "Avarage Number of Steps", type = "l")
```

## Imputing missing values
Here the missing values are replaced by the avarage values within the same time intervall.
```{r}
count <- 0
stepsWithSurrogate <- steps
for(i in 1:length(stepsWithSurrogate$steps)){
  if (is.na(stepsWithSurrogate$steps[i])){
    count <- count + 1
    stepsWithSurrogate$steps[i] <- avgSteps[avgSteps$Group.1 == convertTime(stepsWithSurrogate$interval[i]),2]
    }
}
print(paste("Number of missing steps information: ", count))
stepsPerDaySurrogate <- aggregate(stepsWithSurrogate$steps, by = list(stepsWithSurrogate$date), FUN = sum)
names(stepsPerDaySurrogate) <- c("Date","Steps")
hist(stepsPerDaySurrogate$Steps, main = "Distribution of steps taken per day", xlab = "Steps per Day", breaks = 20)
print(paste("Mean w/o replacemant vs. w/ replacemant: ",mean(stepsPerDay$Steps), mean(stepsPerDaySurrogate$Steps)))
print(paste("Median w/o replacemant vs. w/ replacemant: ",mean(stepsPerDay$Steps), mean(stepsPerDaySurrogate$Steps)))
```

## Are there differences in activity patterns between weekdays and weekends?
Comparison of acitivity at the weekend and on workingdays
```{r}
stepsWeekend <- steps
## Initialize weekend boolean indicator
weekend <- c()
for(i in 1:length(stepsWithSurrogate[,1])){
  weekend <- c(weekend,FALSE)
}
stepsWeekend <- cbind(stepsWeekend, weekend)
stepsWeekend[weekdays(as.Date(stepsWeekend[,2])) == "Sunday", 4] <- TRUE
stepsWeekend[weekdays(as.Date(stepsWeekend[,2])) == "Saturday", 4] <- TRUE
stepsWeekend <- stepsWeekend[!is.na(stepsWeekend$steps),]

avgWeekday <- stepsWeekend[stepsWeekend$weekend == FALSE, ]
avgWeekday <- aggregate(avgWeekday$steps, by = list(avgWeekday$interval), FUN = mean)

avgWeekend <- stepsWeekend[stepsWeekend$weekend == TRUE, ]
avgWeekend <- aggregate(avgWeekend$steps, by = list(avgWeekend$interval), FUN = mean)

par(mfrow = c(2,1))
plot(avgWeekday$Group.1, avgWeekday$x, xlab = "Five Minutes Intervalls [h]", ylab = "Avg #Steps",main = "Avarage Number of Steps on Weekdays", type = "l")
plot(avgWeekend$Group.1, avgWeekend$x, xlab = "Five Minutes Intervalls [h]", ylab = "Avg #Steps", main = "Avarage Number of Steps on Weekend", type = "l")
```